Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2016-11-11 apj.tex
working on Jakob's comments

The quantity $R^F_V = A^F_V/E^F(B-V)$ for an individual supernova is unspecified without knowledge of the intrinsic color.
For a qualitative view, we choose the zeros for $V$-band extinction and color-excess such that the bluest supernova roughly correspond to $E(B-V)=0$,
$A_V=0$, and such that the ridge line is close to constant.
With this ansatz, we
estimate $R^T_V$ per supernova shown in Figure~\ref{avebv:fig}, with the expectations of the
 \citet{1999PASP..111...63F} (hereafter referred to as \citetalias{1999PASP..111...63F})  model for fixed $R^F_V$ overplotted.
With this choice of zeropoints, the ridge line corresponds to $R_V^F\lesssim 2.5$, and even lower values of $R_V^F$ below the ridge line.
We emphasize that the shape of the locus in this plot (though not the relative positions of the points)  and the flatness
of the ridge line is specific to our choice of zeropoints, and that our model does not specify the zeropoints.
A specification of per-supernova $R_V$ requires a model that can identify supernovae that suffer no extrinsic effects,
for example by modeling  a constraining distribution for the extrinsic parameters \citep{2016arXiv160904470M}.
\color{black}

@@

The \citet{1989ApJ...345..245C} dust law can be expressed as a two-parameter linear model
\begin{equation}
A_X = a(X)  A_V + b(X) E(B-V).
\end{equation}
The \citetalias{1999PASP..111...63F} model is approximately described by a linear model with
the above form, with
$a(U,B,V,R,I)=\{0.96,   1.00,   1.00,   0.97,   0.78\}$ and $b(U,B,V,R,I)=\{  1.64,   1.06,   0.16,  -0.40,  -0.52\}$ for the case of
$R^F_V=2.5$
\color{red}
and $A^F_V=0.1$,
for the measurement of the updated
\citet{2007ApJ...663.1187H} SN~Ia template at $B$-band peak as seen through our synthetic passbands.
\color{black}
This linear approximation
is good in the ranges of
\color{red}
 $R^F_V$ and $A^F_V$,
\color{black}
and wavelengths under consideration, in that the values of the components  vary by $<5$\%.

Our model can be expressed as
\begin{equation}
A_X = \frac{\gamma^0_X}{\gamma^0_1-\gamma^0_2}  E_{\gamma^0}(B-V) +  \frac{\gamma^1_X}{\gamma^1_B-\gamma^1_V}  E_{\gamma^1}(B-V).
\end{equation}
\color{red}
Our median-fit values for the $\gamma$ coefficients,
$\frac{\gamma^0_X}{\gamma^0_1-\gamma^0_2}  =\{4.78,   3.88,   2.88,   2.21,   1.59\}$ and
$ \frac{\gamma^1_X}{\gamma^1_B-\gamma^1_V}=\{-3.85,   -4.69,   -5.69,  -5.33,  -4.77\}$,
\color{black}
can expressed in terms of the $a(X)$ and $b(X)$ vectors:
\begin{equation}
\begin{pmatrix}
 \frac{\gamma^0_X}{\gamma^0_B-\gamma^0_V} \\
\frac{\gamma^1_X}{\gamma^1_B-\gamma^1_V} 
\end{pmatrix}=
\begin{pmatrix}
\begin{array}{rr}
1.22 & 2.75 \\
0.81 & -5.50
\end{array}
\end{pmatrix} 
\begin{pmatrix}
a(X) \\
b(X)
\end{pmatrix}+
\begin{pmatrix}
\vec{\epsilon}_a \\
\vec{\epsilon}_b
\end{pmatrix},
\end{equation}
where the two residual vectors are
\color{red}
\begin{align}
\vec{\epsilon}_a &=\{0.14, -0.15, -0.07,  0.04, 0.07\}, \\
\vec{\epsilon}_b & =\{0.12, -0.06, -0.30, 0.30, -0.06\}:
\end{align}
\color{black}
both are perpendicular to $a(X)$ and $b(X)$ and
contribute  3\% and 4\% respectively to the total  length of the $\gamma$ vectors.
This result indicates that
the allowed variation of $UBVRI$ allowed by the \citetalias{1999PASP..111...63F} model and our best-fit model are almost identical.

\color{red}
To help visualize the above result, Figure~\ref{plane:fig} shows two perspectives in $UVI$-space
of the vectors corresponding to $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model,
and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model.
The left plot shows that none of the four vectors are closely aligned with another, while the right plot shows that
the planes defined by our model (solid lines) and the  \citetalias{1999PASP..111...63F} model (dashed lines) are nearly coplanar.
If $R^F_V=2.4$ were constant in our sample, then $A_X = A_V\left(a(X) + \frac{b(X)}{2.4}\right)$;  this combination of $a$ and $b$ is
almost perfectly aligned with $\gamma^0_X/(\gamma^0_B-\gamma^0_V)$, as shown in the dot-dashed line.  
The $\gamma^1$ terms capture the remaining magnitude variations due to the fact that $R^F_V$ is not constant in the sample:
if the $\gamma$'s are due to  \citetalias{1999PASP..111...63F} dust, the only way our analysis could recover
both $\gamma^0$ and $\gamma^1$ is for there to be variation in $R^F_V$.

>>>>

\color{blue}
Our model alone cannot determine $R^T_V = A^T_V/E^T(B-V)$ for individual objects, as $k_0^0$ and $k_1^0$ are unconstrained.
For a qualitative view, we choose zeros  such that the $V$-band extinction and color-excess of the bluest supernova roughly correspond to $E(B-V)=0$,
$A_V=0$, and such that the ridge line 
has close to constant $R^T_V$.
\color{black}
With this ansatz, we
estimate $R^T_V$ per supernova shown in Figure~\ref{avebv:fig}, with the expectations of the
 \citet{1999PASP..111...63F} (hereafter referred to as \citetalias{1999PASP..111...63F})  model for fixed $R^F_V$ overplotted.
With this choice of zeropoints, the ridge line corresponds to $R_V^F\lesssim 2.5$, and even lower values of $R_V^F$ below the ridge line.
We emphasize that the shape of the locus in this plot (though not the relative positions of the points)  and the flatness
of the ridge line is specific to our choice of zeropoints, and that our model does not specify the zeropoints.
A specification of per-supernova $R_V$ requires a model that can identify supernovae that suffer no extrinsic effects,
for example by modeling  a constraining distribution for the extrinsic parameters \citep{2016arXiv160904470M}.
\color{black}

@@

The \citet{1989ApJ...345..245C} dust law can be expressed as a two-parameter linear model
\begin{equation}
A_X = a(X)  A_V + b(X) E(B-V).
\end{equation}
The \citetalias{1999PASP..111...63F} model is approximately described by a linear model with
the above form:
\color{red}
for the case of
$R^F_V=2.5$ and $A^F_V=0.1$,
and the measurement of the updated
\citet{2007ApJ...663.1187H} SN~Ia template at $B$-band peak as seen through our synthetic passbands,
$a(U,B,V,R,I)=\{0.96,   1.00,   1.00,   0.97,   0.78\}$ and $b(U,B,V,R,I)=\{  1.64,   1.06,   0.16,  -0.40,  -0.52\}$ 
\color{black}
This linear approximation
is good in the ranges of
\color{red}
 $R^F_V$ and $A^F_V$,
\color{black}
and wavelengths under consideration, in that the values of the components  vary by $<5$\%.

Our model can be expressed as
\begin{equation}
A_X = \frac{\gamma^0_X}{\gamma^0_1-\gamma^0_2}  E_{\gamma^0}(B-V) +  \frac{\gamma^1_X}{\gamma^1_B-\gamma^1_V}  E_{\gamma^1}(B-V).
\end{equation}
\color{red}
Our median-fit values for the $\gamma$ coefficients,
$\frac{\gamma^0_X}{\gamma^0_B-\gamma^0_V}  =\{4.78,   3.88,   2.88,   2.21,   1.59\}$ and
$ \frac{\gamma^1_X}{\gamma^1_B-\gamma^1_V}=\{-3.85,   -4.69,   -5.69,  -5.33,  -4.77\}$,
\color{black}
can expressed in terms of the $a(X)$ and $b(X)$ vectors:
\begin{equation}
\begin{pmatrix}
 \frac{\gamma^0_X}{\gamma^0_B-\gamma^0_V} \\
\frac{\gamma^1_X}{\gamma^1_B-\gamma^1_V} 
\end{pmatrix}=
\begin{pmatrix}
\begin{array}{rr}
1.22 & 2.75 \\
0.81 & -5.50
\end{array}
\end{pmatrix} 
\begin{pmatrix}
a(X) \\
b(X)
\end{pmatrix}+
\begin{pmatrix}
\vec{\epsilon}_a \\
\vec{\epsilon}_b
\end{pmatrix},
\end{equation}
where the two residual vectors are
\color{red}
\begin{align}
\vec{\epsilon}_a &=\{0.14, -0.15, -0.07,  0.04, 0.07\}, \\
\vec{\epsilon}_b & =\{0.12, -0.06, -0.30, 0.30, -0.06\}:
\end{align}
\color{black}
both are perpendicular to $a(X)$ and $b(X)$ and
contribute  3\% and 4\% respectively to the total  length of the $\gamma$ vectors.
This result indicates that
the allowed variation of $UBVRI$ allowed by the \citetalias{1999PASP..111...63F} model and our best-fit model are almost identical.

\color{red}
To help visualize the above result, Figure~\ref{plane:fig} shows two perspectives in $UVI$-space
of the vectors corresponding to $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model,
and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model.
The left plot shows that none of the four vectors are closely aligned with another, while the right plot shows that
the planes defined by our model (solid lines) and the  \citetalias{1999PASP..111...63F} model (dashed lines) are nearly coplanar.
If $R^F_V=2.4$ were constant in our sample, then $A_X = A_V\left(a(X) + \frac{b(X)}{2.4}\right)$;  this combination of $a$ and $b$ is
almost perfectly aligned with $\gamma^0_X/(\gamma^0_B-\gamma^0_V)$, as shown in the dot-dashed line. 
\color{blue}
Note that the specific directions of $\gamma^0$ and
 $\gamma^1$  within their plane are based on the prior that $k_0$ and $k_1$ are uncorrelated.
\color{black}
The $\gamma^1$ terms capture the remaining magnitude variations due to the fact that $R^F_V$ is not constant in the sample:
if the $\gamma$'s are due to  \citetalias{1999PASP..111...63F} dust, the only way our analysis could recover
both $\gamma^0$ and $\gamma^1$ is for there to be variation in $R^F_V$.
